---
title: "In-class Exercise 8: Geographically Weighted Predictive Modelling"
author: "Dr. Kam Tin Seong<br/>Assoc. Professor of Information Systems(Practice)"
institute: "School of Computing and Information Systems,<br/>Singapore Management University"
date: "3/9/2024"
format: 
  revealjs:
    width: 1600
    height: 900
    show-notes: false
    slide-number: true
    show-slide-number: all
date-format: "DD MMM YYYY"
execute: 
  eval: true
  echo: true
  warning: false
  freeze: true
editor: visual
---

## Installing and Loading R packages

```{r}
pacman::p_load(sf, spdep, GWmodel, SpatialML, 
               tmap, rsample, Metrics, tidyverse,
               knitr, kableExtra)
```

## Preparing Data

### Data import

```{r}
mdata <- read_rds("data/model/mdata.rds") %>%
  st_jitter(amount = .05)
```

---

### Data Sampling

The entire data are split into training and test data sets with 65% and 35% respectively by using *initial_split()* of **rsample** package. rsample is one of the package of tigymodels.

```{r}
#| eval: false
set.seed(1234)
resale_split <- initial_split(mdata, 
                              prop = 6.5/10,)
train_data <- training(resale_split)
test_data <- testing(resale_split)
```

---

### Multicollinearity check

::: panel-tabset

#### Correlation Matrix

```{r}
#| echo: false
#| fig-width: 12
#| fig-height: 8
mdata_nogeo <- mdata %>%
  st_drop_geometry()
ggstatsplot::ggcorrmat(mdata_nogeo[, 2:17])
```

#### The code

```{r}
#| eval: false
mdata_nogeo <- mdata %>%
  st_drop_geometry()
ggstatsplot::ggcorrmat(mdata_nogeo[, 2:17])
```
:::

## Building a non-spatial multiple linear regression

```{r}
#| echo: false
train_data <- read_rds("data/model/train_data.rds")
test_data <- read_rds("data/model/test_data.rds")
```

::: panel-tabset

### The report
```{r}
#| echo: false
price_mlr <- lm(resale_price ~ floor_area_sqm +
                  storey_order + remaining_lease_mths +
                  PROX_CBD + PROX_ELDERLYCARE + PROX_HAWKER +
                  PROX_MRT + PROX_PARK + PROX_MALL + 
                  PROX_SUPERMARKET + WITHIN_350M_KINDERGARTEN +
                  WITHIN_350M_CHILDCARE + WITHIN_350M_BUS +
                  WITHIN_1KM_PRISCH,
                data=train_data)
olsrr::ols_regress(price_mlr)
```

### The code chunk
```{r}
#| eval: false
price_mlr <- lm(resale_price ~ floor_area_sqm +
                  storey_order + remaining_lease_mths +
                  PROX_CBD + PROX_ELDERLYCARE + PROX_HAWKER +
                  PROX_MRT + PROX_PARK + PROX_MALL + 
                  PROX_SUPERMARKET + WITHIN_350M_KINDERGARTEN +
                  WITHIN_350M_CHILDCARE + WITHIN_350M_BUS +
                  WITHIN_1KM_PRISCH,
                data=train_data)
olsrr::ols_regress(price_mlr)
```

:::

---

### Multicollinearity check with VIF

::: panel-tabset

#### VIF
```{r}
vif <- performance::check_collinearity(price_mlr)
kable(vif, 
      caption = "Variance Inflation Factor (VIF) Results") %>%
  kable_styling(font_size = 18) 
```

#### Plotting VIF
```{r}
plot(vif) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```
:::

## Predictive Modelling with MLR

::: panel-tabset

### Computing adaptive bandwidth

```{r}
#| eval: false
bw_adaptive <- bw.gwr(resale_price ~ floor_area_sqm +
                  storey_order + remaining_lease_mths +
                  PROX_CBD + PROX_ELDERLYCARE + PROX_HAWKER +
                  PROX_MRT + PROX_PARK + PROX_MALL + 
                  PROX_SUPERMARKET + WITHIN_350M_KINDERGARTEN +
                  WITHIN_350M_CHILDCARE + WITHIN_350M_BUS +
                  WITHIN_1KM_PRISCH,
                  data=train_data,
                  approach="CV",
                  kernel="gaussian",
                  adaptive=TRUE,
                  longlat=FALSE)
```

```{r}
#| eval: false
write_rds(bw_adaptive, "data/model/bw_adaptive.rds")
```


```{r}
#| echo: false
bw_adaptive <- read_rds("data/model/bw_adaptive.rds")
bw_adaptive
```

### Model calibration

```{r}
#| eval: false
gwr_adaptive <- gwr.basic(formula = resale_price ~
                            floor_area_sqm + storey_order +
                            remaining_lease_mths + PROX_CBD + 
                            PROX_ELDERLYCARE + PROX_HAWKER +
                            PROX_MRT + PROX_PARK + PROX_MALL + 
                            PROX_SUPERMARKET + WITHIN_350M_KINDERGARTEN +
                            WITHIN_350M_CHILDCARE + WITHIN_350M_BUS +
                            WITHIN_1KM_PRISCH,
                          data=train_data,
                          bw=bw_adaptive, 
                          kernel = 'gaussian', 
                          adaptive=TRUE,
                          longlat = FALSE)
```

```{r}
write_rds(gwr_adaptive, "data/model/gwr_adaptive.rds")
```


### Modelling results

```{r}
#| echo: false
gwr_adaptive <- read_rds("data/model/gwr_adaptive.rds")
gwr_adaptive
```
:::

## Predictive Modelling with MLR
### Predicting with test data

::: panel-tabset
### Test data bw 
```{r}
#| eval: false
gwr_bw_test_adaptive <- bw.gwr(resale_price ~ floor_area_sqm +
                  storey_order + remaining_lease_mths +
                  PROX_CBD + PROX_ELDERLYCARE + PROX_HAWKER +
                  PROX_MRT + PROX_PARK + PROX_MALL + 
                  PROX_SUPERMARKET + WITHIN_350M_KINDERGARTEN +
                  WITHIN_350M_CHILDCARE + WITHIN_350M_BUS +
                  WITHIN_1KM_PRISCH,
                  data=test_data,
                  approach="CV",
                  kernel="gaussian",
                  adaptive=TRUE,
                  longlat=FALSE)
```

```{r}
#| echo: false
#| eval: false
write_rds(gwr_bw_test_adaptive,
          "data/model/gwr_bw_test.rds")
```

```{r}
#| echo: false
gwr_bw_test_adaptive <- read_rds(
  "data/model/gwr_bw_test.rds")
```

### Predicting

```{r}
#| eval: false
gwr_pred <- gwr.predict(formula = resale_price ~
                          floor_area_sqm + storey_order +
                          remaining_lease_mths + PROX_CBD + 
                          PROX_ELDERLYCARE + PROX_HAWKER + 
                          PROX_MRT + PROX_PARK + PROX_MALL + 
                          PROX_SUPERMARKET + WITHIN_350M_KINDERGARTEN +
                          WITHIN_350M_CHILDCARE + WITHIN_350M_BUS + 
                          WITHIN_1KM_PRISCH, 
                        data=train_data, 
                        predictdata = test_data, 
                        bw=100, 
                        kernel = 'gaussian', 
                        adaptive=TRUE, 
                        longlat = FALSE)

```
:::

## Predictive Modelling: SpatialML methods

::: panel-tabset

### Preparing coordinate data

```{r}
coords <- st_coordinates(mdata)
coords_train <- st_coordinates(train_data)
coords_test <- st_coordinates(test_data)
```

### Dropping geometry field

```{r}
train_data_nogeom <- train_data %>%
  st_drop_geometry()
```

### Calibrating RF model

```{r}
set.seed(1234)
rf <- ranger(resale_price ~ floor_area_sqm + storey_order + 
               remaining_lease_mths + PROX_CBD + PROX_ELDERLYCARE + 
               PROX_HAWKER + PROX_MRT + PROX_PARK + PROX_MALL + 
               PROX_SUPERMARKET + WITHIN_350M_KINDERGARTEN +
               WITHIN_350M_CHILDCARE + WITHIN_350M_BUS + 
               WITHIN_1KM_PRISCH,
             data=train_data_nogeom)
rf
```

### Calibrating with grf()

```{r}
#| eval: false
set.seed(1234)
gwRF_adaptive <- grf(formula = resale_price ~ floor_area_sqm + 
                       storey_order + remaining_lease_mths + 
                       PROX_CBD + PROX_ELDERLYCARE +
                       PROX_HAWKER + PROX_MRT + PROX_PARK + 
                       PROX_MALL + PROX_SUPERMARKET + 
                       WITHIN_350M_KINDERGARTEN +
                       WITHIN_350M_CHILDCARE + 
                       WITHIN_350M_BUS +
                       WITHIN_1KM_PRISCH,
                     dframe=train_data_nogeom, 
                     bw=55,
                     kernel="adaptive",
                     coords=coords_train)
```

### The model

```{r}
gwRF_adaptive <- read_rds("data/model/gwRF_adaptive.rds")
```

:::


## Predicting by using the test data

::: panel-tabset

### Preparing the test data

```{r}
test_data_nogeom <- cbind(
  test_data, coords_test) %>%
  st_drop_geometry()
```

### Predicting with the test data

```{r}
#| eval: false
gwRF_pred <- predict.grf(gwRF_adaptive, 
                           test_data_nogeom, 
                           x.var.name="X",
                           y.var.name="Y", 
                           local.w=1,
                           global.w=0)
```

### Creating DF

```{r}
GRF_pred <- read_rds("data/model/GRF_pred.rds")
GRF_pred_df <- as.data.frame(GRF_pred)
```

```{r}
test_data_pred <- cbind(test_data, 
                        GRF_pred_df)
```
:::

## Visualising the predicted values

::: panel-tabset

### The plot

```{r}
#| echo: false
ggplot(data = test_data_pred,
       aes(x = GRF_pred,
           y = resale_price)) +
  geom_point()
```

### The code chunk

```{r}
#| eval: false
ggplot(data = test_data_pred,
       aes(x = GRF_pred,
           y = resale_price)) +
  geom_point()
```
:::