---
title: "In-class Exercise 5: Geographically Weighted Statistics - gwModel methods"
date: "12 September 2024"
date-modified: "last-modified"
format: 
  html:
    fontsize: 18px
execute: 
  echo: true
  eval: true
  warning: false
  freeze: true
editor: visual  
---

## Loading the package

In this in-class exercise, sf, spdep, tmap, tidyverse, knitr and GWmodel will be used.

::: {.panel-tabset}

### DIY

Using the step you leanred from previous hands-in, install and load the necessary R packages in R environment.

### The code

```{r}
pacman::p_load(sf, spdep, tmap, tidyverse, knitr, GWmodel)
```

:::

## Preparing the Data

For this in-class exercise, Hunan shapefile and Hunan_2012 data file will be used.

::: {.panel-tabset}

### DIY

Using the steps you learned from previous hands-on, import Hunan shapefile and Hunan_2012 data file into R environment. 


### Importing Hunan shapefile

```{r}
hunan <- st_read(dsn = "data/geospatial", 
                 layer = "Hunan")
```

### Importing Hunan_2012 table

```{r}
hunan2012 <- read_csv("data/aspatial/Hunan_2012.csv")
```

:::

```{r}
hunan <- left_join(hunan, hunan2012) %>%
  select(1:4, 7, 15)
```

```{r}
#| fig-width: 12
#| fig-height: 8
basemap <- tm_shape(hunan) +
  tm_polygons() +
  tm_text("NAME_3", size=0.5)

gdppc <- qtm(hunan, "GDPPC")
tmap_arrange(basemap, gdppc, asp=1, ncol=2)
```

```{r}
wm_q <- poly2nb(hunan, 
                queen=TRUE)
summary(wm_q)
```

```{r}
longitude <- map_dbl(hunan$geometry, ~st_centroid(.x)[[1]])
```

```{r}
latitude <- map_dbl(hunan$geometry, ~st_centroid(.x)[[2]])
```

```{r}
coords <- cbind(longitude, latitude)
```

```{r}
k1 <- knn2nb(knearneigh(coords))
```

```{r}
k1dists <- unlist(nbdists(k1, coords,
                          longlat = TRUE))
```

```{r}
#| fig-width: 8 
#| fig-height: 6
plot(hunan$geometry, border="lightgrey")
plot(wm_q, coords, pch = 19, cex = 0.6, add = TRUE, col= "red")
```

## Working with Geographically Weighted Summary Statistics (GWSS)

```{r}
hunan_sp <- hunan %>%
  as_Spatial()
```

```{r}
gwstat <- gwss(data = hunan_sp,
               vars = "GDPPC",
               bw = 6,
               kernel = "bisquare",
               adaptive = TRUE,
               longlat = T)
```
